---
title: "Monopoly"
author: "Yuki Kitamura"
date: "2024-07-16"
output:
  pdf_document: default
---

This is a simulation of a classic board game, Monopoly. The goal is to find out which spaces on the board get landed on the most.
I will not simulate the entire game. I will simulate only the movement of pieces, and will keep track of which squares the pieces land on.


```{r, error = TRUE}
source("Monopoly Script.R")  # edit with your file name
print(Player)
print(take_turn)
```


# Part 1: Test Cases
Demonstrating that I coded the rules by showing the output of several test cases using preset dice.
The output should be very verbose. It should announce the player roles, where the player moves, what spaces get tallies, if they rolled doubles, etc.

### Test Case 1: Space: Go to Jail
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,4), # Preset the dice rolls 
  verbose = TRUE
)

set.seed(16)
player1 <- Player$new(verbose = TRUE, pos = 24)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:1) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```


### Test Case 2: Chance Card and Doubles Tests: Advance to Go, Reading Railroad, Nearest Railroad, Nearest Utility, No Movement
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,4, 4,3, 1,1, 3,4, 5,3),
  verbose = TRUE
)
set.seed(135)
chance <- CardDeck$new(chancedeck, verbose = TRUE)
community <- CardDeck$new(communitydeck, verbose = TRUE)
player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:4) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```


### Test Case 3: Multiple doubles. Community Chest.
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,3, 2,2, 2,1, 3,1), verbose = TRUE)

player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)
for (i in 1:2) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```


### Test Case 4: Doubles three times. Three turns in jail.
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,3, 3,3, 3,3, 5,6, 5,6, 5,6),
  verbose = TRUE
)

player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:4) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```


### Test Case 5: After going to Jail, player's turn ends immediately. Rolling doubles while in Jail gets player out of jail.
```{r, error = TRUE}
dice <- PresetDice$new(
  rolls = c(3,3, 1,2, 3,3, 3,4),
  verbose = TRUE
)

player1 <- Player$new(verbose = TRUE, pos = 25)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:3) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

print(setNames(monopoly$counts, 1:40))
```


### Test Case 6: 20 Predetermined Turns
```{r, error = TRUE}
# Using preset dices
dice <- PresetDice$new(
  rolls = c(6,4, 5,3, 3,5, 4,4, 4,4, 2,2, 4,3, 4,4, 1,4,
            3,4, 1,2, 3,6, 5,4, 5,5, 1,2, 5,4, 3,3, 6,1,
            1,1, 2,3, 5,5, 5,4, 4,1, 2,2, 2,4),
  verbose = TRUE
)
set.seed(2)
chance <- CardDeck$new(chancedeck, verbose = TRUE)
community <- CardDeck$new(communitydeck, verbose = TRUE)
```


```{r, error = TRUE}
player1 <- Player$new(verbose = TRUE)
monopoly <- SpaceTracker$new(verbose = TRUE)

for (i in 1:20) {
  cat("--- Turn", i,"---\n")
  take_turn(player1, monopoly)
  cat("\n")
}

monopoly$counts
cbind(gameboard, counts = monopoly$counts)
```





# Part 2: 1000 simulated games
The next part is to run 1,000 simulations of a two-player game that lasts 150 turns. This is a total of over 3 hundred thousand tosses of the dice - 1000 games x 150 turns x 2 players + additional rolls if the player gets doubles.
The task is to keep track of where the players land. We ultimately want to build a distribution showing which spaces are most likely to be landed upon. Advance the tokens around the board according to the rules. The special situations involve the cards, jail, and rolling doubles. After 150 turns, reset the game and start over. Simulate 1000 games.
The final output will be two tables of the spaces on the board and their frequencies. 
```{r, message = FALSE}
library(dplyr)
```


```{r, error = TRUE}
## Use non-verbose random dice for Part 2
set.seed(2)
chance <- CardDeck$new(chancedeck, verbose = FALSE)
community <- CardDeck$new(communitydeck, verbose = FALSE)
dice <- RandomDice$new()

player1 <- Player$new(verbose = FALSE)
player2 <- Player$new(verbose = FALSE)
monopoly <- SpaceTracker$new(verbose = FALSE)

for (g in 1:2000) {
  if (g %% 100 == 0) {
    cat("#### SIMULATING GAME", g, "##### \n")
  }
  for (i in 1:150) {
    take_turn(player1, monopoly)
    take_turn(player2, monopoly)
  }
}

print(setNames(monopoly$counts, 1:40))

results <- cbind(gameboard, tally = monopoly$counts)
results <- cbind(results, rel = monopoly$counts/sum(monopoly$counts))
print(results)

arrange(results, desc(tally))
```


```{r}
# set colors for the bar plot
color_vec <- rep(NA, 40)
color_vec[c(2,4)] <- "chocolate4" # mediterranean, baltic
color_vec[c(7,9,10)] <- "lightskyblue" # oriental, vermont, connecticut 
color_vec[c(12,14,15)] <- "darkorchid2" # st charles, states, virgina
color_vec[c(17,19,20)] <- "darkorange" # st james, tennessee, new york
color_vec[c(22,24,25)] <- "red2" # kentucky, indiana, illinois
color_vec[c(27,28,30)] <- "gold1" # atlantic, ventnor, marvin
color_vec[c(32,33,35)] <- "darkgreen" # pacific, n. carolina, pennsylvania
color_vec[c(38,40)] <- "blue3" # park place, boardwalk
color_vec[c(6,16,26,36)] <- "gray14" # railroads
color_vec[c(13,29)] <- "palegreen1" # utilities
color_vec[c(8,23,37)] <- "salmon2" # chance
color_vec[c(3,18,34)] <- "deepskyblue" # community chest 

barplot(monopoly$counts, 
  main = "Barplot of Frequency of Landing for Each Space",
  xlab = "Space Number", ylab = "Number of Landings",
  las = 2, col = color_vec, names.arg = 1:40, cex.names = 0.65)

# this one sets the y-limits so the max count of jail doesn't shrink everything
barplot(monopoly$counts, 
  main = "Barplot of Frequency of Landing for Each Space",
  xlab = "Space Number", ylab = "Number of Landings",
  las = 2, col = color_vec, names.arg = 1:40, cex.names = 0.65,
  ylim = c(0, sort(monopoly$counts, decreasing = TRUE)[2]))
```
